### 2015-04-12 abstract.tex
Intro done.

The largest HPC centers spend considerable time supporting software configurations for thousands of users, but the complexity of software is quickly outpacing the capabilities of existing management tools. Modern HPC applications may require specific versions of many dependency libraries, and they must build with many compilers and MPI versions. The configuration space is exponential in size, and it is infeasible to maintain a single software stack for all users at a site. Existing tools are not sufficient to manage this complexity. We present Spack, a novel package management tool that allows users to build software configurations on demand. Spack provides a novel, recursive specification syntax, and it allows arbitrarily many software and dependency combinations to coexist on the same system. Spack's build mechanism ensures that each package can find its dependencies, regardless of a user's runtime environment settings. Spack brings order to the complexity of HPC software installation management.

>>>>

Large HPC centers spend considerable time supporting software for thousands of users, but the complexity of HPC software is quickly outpacing the capabilities of existing software management tools. Scientific applications require specific versions of many compilers, MPI implementations, and dependency libraries, so using a single, standard software stack is infeasible.  However, managing many configurations is difficult because the configuration space is exponential in size.
%
Spack is a package manager used at Lawrence Livermore National Laboratory (LLNL) to manage software complexity. Spack provides a novel, recursive specification syntax to invoke parametric builds of packages and dependencies.  It allows any number of builds to coexist on the same system, and it ensures that installed packages can find their dependencies, {\it regardless of the environment}. We show through real-world use cases that Spack supports a diverse and demanding user base, bringing order to HPC software chaos.

### 2015-04-12 introduction.tex
Intro done.

\todo{1 page}

\begin{verbatim}
- Why do supercomputing centers need this?

	- specs
		- describe concisely
		- query installed stuff
		- modules require standardized set of names (no order)

	- flexibility (managing combinatoric configuration space)

	- RPATHs instead of (or in addition to) modules
		- pkgs work as built when run
		- spack means there is no module incantation required

	- code teams can build own libs & stacks

	- Contributions:
		1. Flexible mechanism for specifying and querying config space (spec)
		2. Implementation of this in spack
		3. 3 detailed use cases outlining our experiences at LLNL
			a. combinatoric versions
			b. multi-version python installation
			c. site-specific and user-specific policies
\end{verbatim}

>>>>

The Livermore Computing (LC) facility at Lawrence Livermore National Laboratory
(LLNL) supports around 2,500 users on 25 different clusters, ranging 
in size from a 1.6 teraflop, 256-core cluster to to the
20 petaflop, 1.6 million-core Sequoia machine, currently ranked first and
third on the Graph500~\cite{graph500} and Top500~\cite{top500}
lists, respectively.
%
%The massively parallel simulations that run on LC machines support
%U.S. Department of Energy missions across a range of scientific domains, 
%including nuclear physics, material science, climate science, mechanical 
%engineering, geophysics, seismology, and stockpile stewardship.
%
The simulations that run on these machines are very complex; some
depend on specific versions of over 50 dependency libraries.
They require specific compilers, build options and MPI implementations to
achieve the best performance, and users may run several
different such codes in the same environment as part of larger 
scientific workflows.

To support the diverse needs of applications, administrators and application
developers 
frequently build, install, and support many different configurations
of math and physics libraries, as well as other software.
Frequently, applications must be rebuilt to fix bugs and to support
new versions of the OS, MPI, compilers, and other dependencies.
Unfortunately, building scientific software is notoriously complex, with
immature build systems that are difficult to adapt to new
machines~\cite{dubois+:comp-sci-eng,hoste+:pyhpc12,wilson+:corr}.
Worse, the space of required builds grows combinatorically
with each new configuration parameter. As a result, LLNL staff
spend countless hours dealing with build and deployment issues.

Existing package management tools can automate some aspects of the build 
process~\cite{bsdports,digirolamo:smithy,dolstra+:icfp08,dolstra+:lisa04,hashdist,homebrew,hoste+:pyhpc12,macports,thiruvathukal:gentoo04}.
For the most part, these systems focus on keeping a single, stable set of 
packages up to date, and they do not handle installation of multiple
versions or configurations.  Those that {\it do} handle multiple configurations
typically require that package files be created for each combination of 
options~ \cite{digirolamo:smithy,dolstra+:icfp08,dolstra+:lisa04,hashdist,hoste+:pyhpc12}, 
leading to a profusion of files and maintenance issues.
To our knowledge, no existing tool allows new configurations to be composed
and assembled on demand.  Some allow limited forms of 
composition~\cite{hoste+:pyhpc12,dolstra+:icfp08,dolstra+:lisa04}, but they
have limited dependency management capabilities, and they still burden
users with combinatoric naming and version management.

In this paper, we describe our experiences with the {\it Spack} package manager,
which we have developed at LLNL to manage increasing software complexity.
Specifically, this paper offers the following contributions:
\begin{enumerate}
\item A novel, recursive syntax for specifying, building, and querying
      HPC packages in a large parameter space;
\item A build methodology that ensures packages find their dependencies
      regardless of users' environment; 
\item Spack: An implementation of these concepts; and
\item Three use cases detailing LLNL's use of Spack to rapidly compose
      software configurations, manage Python installations, and implement
      site-specific build policies.
\end{enumerate}

Spack is in active development at LLNL, and our use cases outline some of
its first uses in our environment.  Spack solves a problem that is pervasive
at large, multi-user HPC centers, and we believe that spack and our
experiences with it can be applied at other HPC facilities to improve
operational efficiency by simplifying the build and deployment of 
bleeding-edge scientific software.
